---
title: "Co-Culture Assay"
author: "Sachida Pokhrel"
date: "2025-04-18"
output: 
  md_document:
    variant: gfm
  html_document:
    toc: TRUE
    toc_float: TRUE
---
## Things to know before Analysis

***This is applied for all the dataset for Co-culture Assay***

- One .csv file is loaded for one bacteria interacting with multiple yeast isolates
- There are three R chunks for each bacterial data. 
- There are 16 different yeast interacting with one bacteria. So, to make it more clear in the plot and to make it less cumbersome, these yeast isolates are further categorised into 4 Yeast classes. 
- There are 8 colonies of yeast and 8 colonies of bacteria at different distances in one plate. Data is taken only for bacterial colony size which reflects the impact of yeast on bacteria.
- Plots have multiple line graphs, thus the significant letters are not assigned through code and will be added later after exporting the plot.

```{r Load necessary libraries}
#load necessary library
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(gvlma)
library(agricolae)
library(emmeans)
library(multcompView)
library(dplyr)
library(stringr)
library(tidyverse)
library(car)
library(nlme)
```

## ***Curtobacterium flaccumfaciens* EMM_B44**
Plot is generated using loop around the 4 different classes of yeast, coming up with 4 plots as an output which will be combined in one plot.

```{r Plot for EMM_B44}
#read data
B44 <- read.csv("CoCultureAssay/CoCultureAssayData/2024-08-09_PeaceAssay_B44.csv", na.strings = "na") 


#load cbb color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
str(B44)
#round off of the distance to yeast so that it is visibly a categorical variable
B44$distance_to_yeast <- round(B44$distance_to_yeast, 0)
#change into categorical dataset
B44$Replication=as.factor(B44$Replication)
B44$Yeast=as.factor(B44$Yeast)
B44$DAI=as.factor(B44$DAI)
B44$distance_to_yeast=as.factor(B44$distance_to_yeast)

#remove the data for the distance between yeast and bacteria at zero to remove misleading details
B44.no.contact <- B44[B44$distance_to_yeast != 0, ]
head(B44.no.contact)

#remove first data since it is a reference for increase in colony size and donot follow the assumption of linear model i.e. normal distributuion
B44.no.1st.data <-B44.no.contact[B44.no.contact$DAI != "2",]

# For plots purpose, I am only taking Day 6 since it has visible difference compared to other 
B44.Day6 <- B44.no.contact[B44.no.contact$DAI == "6",]

#making a list object to store all the plots
plot_listB44 <- list()

# Get unique classes from the data
unique_class <- unique(B44.Day6$Class[B44.Day6$Class != "Control"])

# Loop through each class
for (i in seq_along(unique_class)) {
  
  # Filter for Control and the current class for using it in one iteration
  data <- B44.Day6 %>%
    filter(Class %in% c("Control", unique_class[i]))
  
  # Generate the plot
  p <- ggplot(data, aes(x = distance_to_yeast, y = increase, group = Yeast, color = Yeast)) +
    stat_summary(fun = mean, geom = "line") +   #making a line graph
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +   #standard error in plot
    ylab("Increase in colony diameter (mm)") +   
    xlab("Distance to Yeast (mm)") +   
    theme_classic() +
    scale_color_manual(values = cbbPalette) + #providing color blind palette
    ggtitle(unique_class[i]) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Store the plot in the list
  plot_listB44[[i]] <- p
}

#combining each iteration into one plot for one bacteria or dataset
combined_plotB44 <- ggarrange(plotlist = plot_listB44, ncol = 2, nrow = 2, common.legend = FALSE) #using the list generated from the loop, use it to make a combined plot without common legend.
# Annotation for the title
final_plotB44 <- annotate_figure(combined_plotB44,
                                 top = text_grob(
    expression("Impact on growth of"~italic("Curtobacterium flaccumfaciens")~"EMM_B44 by Yeast"), color = "Blue2", face = "bold", size = 14, hjust = 0.5))

#display final plot
print(final_plotB44)

```

### Stats for *Curtobacterium flaccumfaciens* EMM_B44
We are using linear mixed model. Our dependent variable or y is increase (increase in colony diameter from 1st data) and independent variables are different Yeast isolates, days after inoculation (DAI), and distance to yeast which is the distance between the yeast and bacterial colony in the plate. Each plate is replicated 3 times. 

***This applies for all the dataset for Co-Culture Plate Assay***

```{r Stats for EMM_B44}
#filter data to remove 1st day data since the first data is taken as a base to measure the increase in colony size to rule out the variability that is caused by the drop inoculation. So, initially the increase in the colony diameter for 1st data for all colony is "0" that violates the assumption of normality, thus we remove that from analysis. This would be similar for all the bacterial isolates.

B44.no.1st.data <- B44.no.contact[B44.no.contact$DAI != "2",]
#We check interaction for all before cutting it down to the significant one.
B44try=lme(increase~DAI*distance_to_yeast*Yeast, data = B44.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(B44try)

#removed interaction terms that were not significant.
resultsB44=lme(increase~DAI+distance_to_yeast*Yeast, data = B44.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(B44try, resultsB44) #check if complex model was better or not
summary(resultsB44) #gives summary of the linear model
anova(resultsB44) #for overall x variables
```

### Loop for running analysis for each day separately for EMM_B44
Days after inoculation (DAI) as a factor is always significantly impacting the growth. Also, our plot will represent data for Day 6 thus we want relevant stats and comparison on Day 6 to present in the plot. So, loop was made for each Day data and removing DAI from the model and keeping rest of it present. 
```{r loop of stat for each day for EMM_B44}
# Extract all the unique days to be used for iteration in for loop
day <- unique(B44.no.1st.data$DAI)
# Initialize empty lists to store results
B44_Day <- list()
meansB44_Day <- list()
resultsB44_Day <- list()
model_summary_B44 <- list()
anova_table_B44 <- list()
Results_lsmeansB44 <- list()
filtered_comparisonsB44 <- list()

for (i in seq_along(day)) { #there are DAI 4, 6 and 8 that was extracted as unique days; sequence assigned will be 1 , 2 and 3 for each unique DAI
  
   # Filter data for the current day
  B44_Day[[i]] <- filter(B44.no.1st.data, DAI == day[i])
    # Fit the mixed effects model
  resultsB44_Day[[i]] <- lme(increase ~ distance_to_yeast * Yeast,
                            data = B44_Day[[i]],
                            random = ~1 | Replication,
                            na.action = na.omit) #Using mixed effect model here subsetting the data for the particular day

    model_summary_B44[[i]] <- summary(resultsB44_Day[[i]]) #store in the list value
  anova_table_B44[[i]] <- anova(resultsB44_Day[[i]]) #store in list value
   # Estimated marginal means
  meansB44_Day[[i]] <- emmeans(resultsB44_Day[[i]], ~ Yeast | distance_to_yeast)
    # Compact letter display with Tukey adjustment
  Results_lsmeansB44[[i]] <- multcomp::cld(meansB44_Day[[i]], alpha = 0.05,
                                          Letters = letters, reversed = TRUE,
                                          details = TRUE)
   # Print lsmeans results
  #print(Results_lsmeansB44[[i]])
  
  # Filter comparisons involving "Control"
  filtered_comparisonsB44[[i]] <- Results_lsmeansB44[[i]]$comparisons %>%
    filter(grepl("Control", contrast))
   # Print filtered contrasts
  print(filtered_comparisonsB44[[i]])
}

```


## ***Curtobacterium flaccumfaciens* EMM_B30**

```{r Plot for EMM_B30}
#read data
B30 <- read.csv("CoCultureAssay/CoCultureAssayData/2024-07-21_PeaceAssay_B30.csv", na.strings = "na") 


#load cbb color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
str(B30)
#round off of the distance to yeast so that it is visibly a categorical variable
B30$distance_to_yeast <- round(B30$distance_to_yeast, 0)
#change into categorical dataset
B30$Replication=as.factor(B30$Replication)
B30$Yeast=as.factor(B30$Yeast)
B30$DAI=as.factor(B30$DAI)
B30$distance_to_yeast=as.factor(B30$distance_to_yeast)

#remove the data for the distance zero
B30.no.contact <- B30[B30$distance_to_yeast != 0, ]
head(B30.no.contact)

# remove data first data since it is a reference for increase in colony size and donot follow the assumption of linear model i.e. normal distribution
B30.no.1st.data <-B30.no.contact[B30.no.contact$DAI != "2",]


B30.Day6 <- B30.no.contact[B30.no.contact$DAI == "6",]


plot_listB30 <- list()

# Get unique classes
unique_class <- unique(B30.Day6$Class[B30.Day6$Class != "Control"])

# Loop through each class
for (i in seq_along(unique_class)) {
  
  # Filter for Control and the current class
  data <- B30.Day6 %>%
    filter(Class %in% c("Control", unique_class[i]))
  
  # Generate the plot
  p <- ggplot(data, aes(x = distance_to_yeast, y = increase, group = Yeast, color = Yeast)) +
    stat_summary(fun = mean, geom = "line") +   
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +   
    ylab("Increase in colony diameter (mm)") +   
    xlab("Distance to Yeast (mm)") +   
    theme_classic() +
    scale_color_manual(values = cbbPalette) +
    ggtitle(unique_class[i]) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Store the plot
  plot_listB30[[i]] <- p
}

combined_plotB30 <- ggarrange(plotlist = plot_listB30, ncol = 2, nrow = 2, common.legend = FALSE)
# Annotation for the title
final_plotB30 <- annotate_figure(combined_plotB30,
                               top = text_grob(
    expression("Impact on growth of"~italic("Curtobacterium flaccumfaciens")~"EMM_B30 by Yeast"),, color = "Blue2", face = "bold", size = 14, hjust = 0.5))

print(final_plotB30)
```

***After first bacteria everything is repeated in similar way for all dataset***

### Stats for *Curtobacterium flaccumfaciens* EMM_B30
```{r Stats for EMM_B30}
B30.no.1st.data <- B30.no.contact[B30.no.contact$DAI != "2",]
B30try <- lme(increase~DAI*distance_to_yeast*Yeast, data = B30.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(B30try)

resultsB30=lme(increase~DAI+distance_to_yeast*Yeast, data = B30.no.1st.data, random = ~1|Replication, na.action = na.omit)
summary(resultsB30)
#Anova(resultsB5)
anova(resultsB30)
```

### Loop for running analysis for each day separately for EMM_B30
```{r Loop of stat for each day for EMM_B30}
# Unique days
day <- unique(B30.no.1st.data$DAI)
# Initialize empty lists to store results
B30_Day <- list()
meansB30_Day <- list()
resultsB30_Day <- list()
model_summary_B30 <- list()
anova_table_B30 <- list()
Results_lsmeansB30 <- list()
filtered_comparisonsB30 <- list()

for (i in seq_along(day)) {
    # Filter data for the current day
  B30_Day[[i]] <- filter(B30.no.1st.data, DAI == day[i])
    # Fit the mixed effects model
  resultsB30_Day[[i]] <- lme(increase ~ distance_to_yeast * Yeast,
                            data = B30_Day[[i]],
                            random = ~1 | Replication,
                            na.action = na.omit)
    model_summary_B30[[i]] <- summary(resultsB30_Day[[i]])
  anova_table_B30[[i]] <- anova(resultsB30_Day[[i]])
   # Estimated marginal means
  meansB30_Day[[i]] <- emmeans(resultsB30_Day[[i]], ~ Yeast | distance_to_yeast)
    # Compact letter display with Tukey adjustment
  Results_lsmeansB30[[i]] <- multcomp::cld(meansB30_Day[[i]], alpha = 0.05,
                                          Letters = letters, reversed = TRUE,
                                          details = TRUE)
   # Print lsmeans results
  print(Results_lsmeansB30[[i]])
  
  # Filter comparisons involving "Control"
  filtered_comparisonsB30[[i]] <- Results_lsmeansB30[[i]]$comparisons %>%
    filter(grepl("Control", contrast))
   # Print filtered contrasts
  print(filtered_comparisonsB30[[i]])
}
```

## ***Pseudomonas massiliensis* EMM_B5**

```{r Plot for EMM_B5}
#read data
B5 <- read.csv("CoCultureAssay/CoCultureAssayData/2024-08-09_PeaceAssay_B5.csv", na.strings = "na") 


#load cbb color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
str(B5)
#round off of the distance to yeast so that it is visibly a categorical variable
B5$distance_to_yeast <- round(B5$distance_to_yeast, 0)
#change into categorical dataset
B5$Replication=as.factor(B5$Replication)
B5$Yeast=as.factor(B5$Yeast)
B5$DAI=as.factor(B5$DAI)
B5$distance_to_yeast=as.factor(B5$distance_to_yeast)

#remove the data for the distance zero
B5.no.contact <- B5[B5$distance_to_yeast != 0, ]
head(B5.no.contact)

# remove data first data since it is a reference for increase in colony size and donot follow the assumption of linear model i.e. normal distributuion
B5.Day6 <- B5.no.contact[B5.no.contact$DAI == "6",]


plot_listB5 <- list()

# Get unique classes
unique_class <- unique(B5.Day6$Class[B5.Day6$Class != "Control"])

# Loop through each class
for (i in seq_along(unique_class)) {
  
  # Filter for Control and the current class
  data <- B5.Day6 %>%
    filter(Class %in% c("Control", unique_class[i]))
  
  # Generate the plot
  p <- ggplot(data, aes(x = distance_to_yeast, y = increase, group = Yeast, color = Yeast)) +
    stat_summary(fun = mean, geom = "line") +   
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +   
    ylab("Increase in colony diameter (mm)") +   
    xlab("Distance to Yeast (mm)") +   
    theme_classic() +
    scale_color_manual(values = cbbPalette) +
    ggtitle(unique_class[i]) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Store the plot
  plot_listB5[[i]] <- p
}

combined_plotB5 <- ggarrange(plotlist = plot_listB5, ncol = 2, nrow = 2, common.legend = FALSE)

# Annotation for the title
final_plotB5 <- annotate_figure(combined_plotB5,
                              top = text_grob(
    expression("Impact on growth of"~italic("Pseudomonas massiliensis")~"EMM_B5 by Yeast"), color = "Blue2", face = "bold", size = 14, hjust = 0.5))

print(final_plotB5)

```

### Stats *Pseudomonas massiliensis* EMM_B5

```{r Stats for EMM_B5}
library(nlme)
B5.no.1st.data <- B5.no.contact[B5.no.contact$DAI != "2",]
B5try <- lme(increase~distance_to_yeast*Yeast*DAI, data = B5.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(B5try)

resultsB5=lme(increase~distance_to_yeast*Yeast+Yeast*DAI, data = B5.no.1st.data, random = ~1|Replication, na.action = na.omit)
summary(resultsB5)
```

### Loop for running analysis for each day separately EMM_B5

```{r Loop of stat for each day for EMM_B5}
# Unique days
day <- unique(B5.no.1st.data$DAI)
# Initialize empty lists to store results
B5_Day <- list()
meansB5_Day <- list()
resultsB5_Day <- list()
model_summary_B5 <- list()
anova_table_B5 <- list()
Results_lsmeansB5 <- list()
filtered_comparisonsB5 <- list()

for (i in seq_along(day)) {
    # Filter data for the current day
  B5_Day[[i]] <- filter(B5.no.contact, DAI == day[i])
    # Fit the mixed effects model
  resultsB5_Day[[i]] <- lme(increase ~ distance_to_yeast*Yeast,
                            data = B5_Day[[i]],
                            random = ~1 | Replication,
                            na.action = na.omit)
    model_summary_B5[[i]] <- summary(resultsB5_Day[[i]])
  anova_table_B5[[i]] <- anova(resultsB5_Day[[i]])
   # Estimated marginal means
  meansB5_Day[[i]] <- emmeans(resultsB5_Day[[i]], ~ Yeast|distance_to_yeast)
    # Compact letter display with Tukey adjustment
  Results_lsmeansB5[[i]] <- multcomp::cld(meansB5_Day[[i]], alpha = 0.05,
                                          Letters = letters, reversed = TRUE,
                                          details = TRUE)
   # Print lsmeans results
 # print(Results_lsmeansB5[[i]])
  
  # Filter comparisons involving "Control"
  filtered_comparisonsB5[[i]] <- Results_lsmeansB5[[i]]$comparisons %>%
    filter(grepl("Control", contrast))
   # Print filtered contrasts
  print(filtered_comparisonsB5[[i]])
}

```

## ***Xanthomonas perforans* AL65**

```{r Plot for AL65}
#read data
AL65 <- read.csv("CoCultureAssay/CoCultureAssayData/2024-08-09_PeaceAssay_AL65.csv", na.strings = "na") 
#load cbb color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
str(AL65)
#round off of the distance to yeast so that it is visibly a categorical variable
AL65$distance_to_yeast <- round(AL65$distance_to_yeast, 0)
#change into categorical dataset
AL65$Replication=as.factor(AL65$Replication)
AL65$Yeast=as.factor(AL65$Yeast)
AL65$DAI=as.factor(AL65$DAI)
AL65$distance_to_yeast=as.factor(AL65$distance_to_yeast)

#remove the data for the distance zero
AL65.no.contact <- AL65[AL65$distance_to_yeast != 0, ]
head(AL65.no.contact)

# remove data first data since it is a reference for increase in colony size and donot follow the assumption of linear model i.e. normal distribution
AL65.no.1st.data <- AL65.no.contact[AL65.no.contact$DAI != "2",]

#Using Day 6th data for plot
AL65.Day6 <- AL65.no.contact[AL65.no.contact$DAI == "6",]


plot_listAL65 <- list()

# Get unique classes
unique_class <- unique(AL65.Day6$Class[AL65.Day6$Class != "Control"])

# Loop through each class
for (i in seq_along(unique_class)) {
  
  # Filter for Control and the current class
  data <- AL65.Day6 %>%
    filter(Class %in% c("Control", unique_class[i]))
  
  # Generate the plot
  p <- ggplot(data, aes(x = distance_to_yeast, y = increase, group = Yeast, color = Yeast)) +
    stat_summary(fun = mean, geom = "line") +   
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +   
    ylab("Increase in colony diameter (mm)") +   
    xlab("Distance to Yeast (mm)") +   
    theme_classic() +
    scale_color_manual(values = cbbPalette) +
    ggtitle(unique_class[i]) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Store the plot
  plot_listAL65[[i]] <- p
}

combined_plotAL65 <- ggarrange(plotlist = plot_listAL65, ncol = 2, nrow = 2, common.legend = FALSE)

# Annotation for the title
final_plotAL65 <- annotate_figure(combined_plotAL65,
                               top = text_grob(
    expression("Impact on growth of"~italic("Xanthomonas perforans")~"AL65 by Yeast"), face = "bold", color = "Blue2", size = 14, hjust = 0.5))

print(final_plotAL65)

```



### Stats for *Xanthomonas perforans* AL65

```{r Stats for AL65}
AL65.no.1st.data <- AL65.no.contact[AL65.no.contact$DAI != "2",]
AL65try <- lme(increase~DAI*distance_to_yeast*Yeast, data = AL65.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(AL65try)

resultsAL65=lme(increase~DAI*distance_to_yeast+Yeast, data = AL65.no.1st.data, random = ~1|Replication, na.action = na.omit)
summary(resultsAL65)
anova(resultsAL65)
```

### Loop for running analysis for each day separately for AL65
```{r Loop of stat for each day for AL65}
# Unique days
day <- unique(AL65.no.1st.data$DAI)
# Initialize empty lists to store results
AL65_Day <- list()
meansAL65_Day <- list()
resultsAL65_Day <- list()
model_summary_AL65 <- list()
anova_table_AL65 <- list()
Results_lsmeansAL65 <- list()
filtered_comparisonsAL65 <- list()

for (i in seq_along(day)) {
    # Filter data for the current day
  AL65_Day[[i]] <- filter(AL65.no.1st.data, DAI == day[i])
    # Fit the mixed effects model
  resultsAL65_Day[[i]] <- lme(increase ~ distance_to_yeast+Yeast,
                            data = AL65_Day[[i]],
                            random = ~1 | Replication,
                            na.action = na.omit)
    model_summary_AL65[[i]] <- summary(resultsAL65_Day[[i]])
  anova_table_AL65[[i]] <- anova(resultsAL65_Day[[i]])
   # Estimated marginal means
  meansAL65_Day[[i]] <- emmeans(resultsAL65_Day[[i]], ~ Yeast | distance_to_yeast)
    # Compact letter display with Tukey adjustment
  Results_lsmeansAL65[[i]] <- multcomp::cld(meansAL65_Day[[i]], alpha = 0.05,
                                          Letters = letters, reversed = TRUE,
                                          details = TRUE)
   # Print lsmeans results
  #print(Results_lsmeansAL65[[i]])
  
  # Filter comparisons involving "Control"
  filtered_comparisonsAL65[[i]] <- Results_lsmeansAL65[[i]]$comparisons %>%
    filter(grepl("Control", contrast))
   # Print filtered contrasts
  print(filtered_comparisonsAL65[[i]])
}

```


## ***Hymenobacter polaris* EMM_B17**

```{r Plot for EMM_B17}
#read data
B17 <- read.csv("CoCultureAssay/CoCultureAssayData/2024-07-17_PeaceAssay_B17.csv", na.strings = "na") 


#load cbb color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
str(B17)
#round off of the distance to yeast so that it is visibly a categorical variable
B17$distance_to_yeast <- round(B17$distance_to_yeast, 0)
#change into categorical dataset
B17$Replication=as.factor(B17$Replication)
B17$Yeast=as.factor(B17$Yeast)
B17$DAI=as.factor(B17$DAI)
B17$distance_to_yeast=as.factor(B17$distance_to_yeast)

#remove the data for the distance zero
B17.no.contact <- B17[B17$distance_to_yeast != 0, ]
head(B17.no.contact)

# remove data first data since it is a reference for increase in colony size and donot follow the assumption of linear model i.e. normal distributuion
B17.Day6 <- B17.no.contact[B17.no.contact$DAI == "6",]


plot_listB17 <- list()

# Get unique classes
unique_class <- unique(B17.Day6$Class[B17.Day6$Class != "Control"])

# Loop through each class
for (i in seq_along(unique_class)) {
  
  # Filter for Control and the current class
  data <- B17.Day6 %>%
    filter(Class %in% c("Control", unique_class[i]))
  
  # Generate the plot
  p <- ggplot(data, aes(x = distance_to_yeast, y = increase, group = Yeast, color = Yeast)) +
    stat_summary(fun = mean, geom = "line") +   
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +   
    ylab("Increase in colony diameter (mm)") +   
    xlab("Distance to Yeast (mm)") +   
    theme_classic() +
    scale_color_manual(values = cbbPalette) +
    ggtitle(unique_class[i]) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Store the plot
  plot_listB17[[i]] <- p
}

combined_plotB17 <- ggarrange(plotlist = plot_listB17, ncol = 2, nrow = 2, common.legend = FALSE)

# Annotation for the title
final_plotB17 <- annotate_figure(combined_plotB17,
                                top = text_grob(
    expression("Impact on growth of"~italic("Hymenobacter polaris")~"EMM_B17 by Yeast"), face = "bold", color = "Blue2", size = 14, hjust = 0.5))

print(final_plotB17)
```

### Stats for *Hymenobacter polaris* EMM_B17
```{r Stats for EMM_B17}
B17.no.1st.data <- B17.no.contact[B17.no.contact$DAI != "2",]
B17try <- lme(increase~DAI*distance_to_yeast*Yeast, data = B17.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(B17try)

resultsB17=lme(increase~DAI+distance_to_yeast*Yeast, data = B17.no.1st.data, random = ~1|Replication, na.action = na.omit)
summary(resultsB17)
#Anova(resultsB5)
anova(resultsB17)
```

### Loop for running analysis for each day separately for EMM_B17

```{r Loop of stat for each day for EMM_B17}
# Unique days
day <- unique(B17.no.1st.data$DAI)
# Initialize empty lists to store results
B17_Day <- list()
meansB17_Day <- list()
resultsB17_Day <- list()
model_summary_B17 <- list()
anova_table_B17 <- list()
Results_lsmeansB17 <- list()
filtered_comparisonsB17 <- list()

for (i in seq_along(day)) {
    # Filter data for the current day
  B17_Day[[i]] <- filter(B17.no.1st.data, DAI == day[i])
    # Fit the mixed effects model
  resultsB17_Day[[i]] <- lme(increase ~ distance_to_yeast * Yeast,
                            data = B17_Day[[i]],
                            random = ~1 | Replication,
                            na.action = na.omit)
    model_summary_B17[[i]] <- summary(resultsB17_Day[[i]])
  anova_table_B17[[i]] <- anova(resultsB17_Day[[i]])
   # Estimated marginal means
  meansB17_Day[[i]] <- emmeans(resultsB17_Day[[i]], ~ Yeast | distance_to_yeast)
    # Compact letter display with Tukey adjustment
  Results_lsmeansB17[[i]] <- multcomp::cld(meansB17_Day[[i]], alpha = 0.05,
                                          Letters = letters, reversed = TRUE,
                                          details = TRUE)
   # Print lsmeans results
  #print(Results_lsmeansB17[[i]])
  
  # Filter comparisons involving "Control"
  filtered_comparisonsB17[[i]] <- Results_lsmeansB17[[i]]$comparisons %>%
    filter(grepl("Control", contrast))
   # Print filtered contrasts
  print(filtered_comparisonsB17[[i]])
}

```


## ***Methylobacterium platani* EMM_B52**

```{r Plot for EMM_B52}
#read data
B52 <- read.csv("CoCultureAssay/CoCultureAssayData/MergedB52.csv", na.strings = "na") 
#load cbb color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
str(B52)
#round off of the distance to yeast so that it is visibly a categorical variable
B52$distance_to_yeast <- round(B52$distance_to_yeast, 0)
#change into categorical dataset
B52$Replication=as.factor(B52$Replication)
B52$Yeast=as.factor(B52$Yeast)
B52$DAI=as.factor(B52$DAI)
B52$distance_to_yeast=as.factor(B52$distance_to_yeast)

#remove the data for the distance zero
B52.no.contact <- B52[B52$distance_to_yeast != 0, ]
head(B52.no.contact)

# remove data first data since it is a reference for increase in colony size and donot follow the assumption of linear model i.e. normal distribution
B52.no.1st.data <- B52.no.contact[B52.no.contact$DAI != "2",]

#for plot purposes we used Day 6 data
B52.Day6 <- B52.no.contact[B52.no.contact$DAI == "6",]


plot_listB52 <- list()

# Get unique classes
unique_class <- unique(B52.Day6$Class[B52.Day6$Class != "Control"])

# Loop through each class
for (i in seq_along(unique_class)) {
  
  # Filter for Control and the current class
  data <- B52.Day6 %>%
    filter(Class %in% c("Control", unique_class[i]))
  
  # Generate the plot
  p <- ggplot(data, aes(x = distance_to_yeast, y = increase, group = Yeast, color = Yeast)) +
    stat_summary(fun = mean, geom = "line") +   
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +   
    ylab("Increase in colony diameter (mm)") +   
    xlab("Distance to Yeast (mm)") +   
    theme_classic() +
    scale_color_manual(values = cbbPalette) +
    ggtitle(unique_class[i]) +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Store the plot
  plot_listB52[[i]] <- p
}

combined_plotB52 <- ggarrange(plotlist = plot_listB52, ncol = 2, nrow = 2, common.legend = FALSE)

# Annotation for the title
final_plotB52 <- annotate_figure(combined_plotB52,
                                  top = text_grob(
    expression("Impact on growth of"~italic("Methylobacterium platanii")~"EMM_B52 by Yeast"), face = "bold", color = "Blue2", size = 14, hjust = 0.5))

print(final_plotB52)
```

### Stats *Methylobacterium platanii* EMM_B52

```{r Stats for EMM_B52}
B52.no.1st.data <- B52.no.contact[B52.no.contact$DAI != "2",]
B52try <- lme(increase~distance_to_yeast*DAI*Yeast, data = B52.no.1st.data, random = ~1|Replication, na.action = na.omit)
anova(B52try)

resultsB52=lme(increase~DAI+distance_to_yeast*Yeast+DAI*Yeast, data = B52.no.1st.data, random = ~1|Replication, na.action = na.omit)
summary(resultsB52)
#Anova(resultsB5)
anova(resultsB52)
```

### Loop for running analysis for each day separately for EMM_B52
```{r Loop of stat for each day for EMM_B44}
# Unique days
day <- unique(B52.no.1st.data$DAI)
# Initialize empty lists to store results
B52_Day <- list()
meansB52_Day <- list()
resultsB52_Day <- list()
model_summary_B52 <- list()
anova_table_B52 <- list()
Results_lsmeansB52 <- list()
filtered_comparisonsB52 <- list()

for (i in seq_along(day)) {
  # Filter data for the current day
  B52_Day[[i]] <- filter(B52.no.1st.data, DAI == day[i])
  # Fit the mixed effects model
  resultsB52_Day[[i]] <- lme(increase ~ distance_to_yeast * Yeast,
                             data = B52_Day[[i]],
                             random = ~1 | Replication,
                             na.action = na.omit)
  model_summary_B52[[i]] <- summary(resultsB52_Day[[i]])
  anova_table_B52[[i]] <- anova(resultsB52_Day[[i]])
  # Estimated marginal means
  meansB52_Day[[i]] <- emmeans(resultsB52_Day[[i]], ~ Yeast | distance_to_yeast)
  # Compact letter display with Tukey adjustment
  Results_lsmeansB52[[i]] <- multcomp::cld(meansB52_Day[[i]], alpha = 0.05,
                                           Letters = letters, reversed = TRUE,
                                           details = TRUE)
  # Print lsmeans results
  #print(Results_lsmeansB52[[i]])
  
  # Filter comparisons involving "Control"
  filtered_comparisonsB52[[i]] <- Results_lsmeansB52[[i]]$comparisons %>%
    filter(grepl("Control", contrast))
  # Print filtered contrasts
  print(filtered_comparisonsB52[[i]])
}

```
 


