---
title: "Split-Plate Assay"
author: "Sachida Pokhrel"
date: "2025-04-16"
output: 
  md_document:
    variant: gfm
  html_document:
    toc: TRUE
    toc_float: TRUE
---

## Things to know before analysis

***This is applied to both replicate***

- Independent variable is colony diameter, colony weight and colony forming Unit
- Dependent variable is DAI (Days after Inoculation) in case of diameter, yeast isolates for all parameters


```{r Load necessary libraries}
#load the library
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(emmeans)
library(multcompView)
library(dplyr)
library(stringr)
library(tidyverse)
library(car)
```

## Biological Replicate 1
```{r Read data for Biological Replicate 1}
#read data
BioRep1 <- read.csv("Splitplate/SplitPlateData/2024-09-16_SplitPlate_B52.csv", na.strings = "na")
#load the color palettes
cbbPalette <- c( "#E69F00", "#56B4E9", "#009E73", "#CC79A7")
#check the structure and see the data
str(BioRep1)
head(BioRep1)
#changing the variables into categorical data
BioRep1$Replication=as.factor(BioRep1$Replication)
BioRep1$Media=as.factor(BioRep1$Media)
BioRep1$Yeast=as.factor(BioRep1$Yeast)
BioRep1$DAI=as.factor(BioRep1$DAI)
```

### Impact of Yeast on Colony Size

```{r Increase in colony size Plot for Biological replicate 1}
#filter 1st data date and removing the 1st data since there was no growth associated with it. The data is considered to be the baseline for growth so since it violates the assumption of normality, we remove it from the analysis

filterdata1 <-BioRep1[BioRep1$Media == "YePDA"& BioRep1$DAI != 2,]
#load library
library(nlme)

#stats used for size
size1=lme(increase~Yeast*DAI, data = filterdata1,random=~1|Replication) #mixed effect model
size2=lm(increase ~ Yeast*DAI, data = filterdata1) #use simpler model
summary(size1)
anova(size1, size2) #to check if complex or simple model is better

#pairwise comparison
lsmeans.size1 <- emmeans(size1, ~ Yeast|DAI) 
# estimate lsmeans 
comp.size1 <- multcomp::cld(object = lsmeans.size1, alpha = 0.05, Letters = letters, reversed = TRUE, details = TRUE)
# contrast with Tukey ajustment 
comp.size1


# Extracting the letters for the bars
sig.diff.letters <- data.frame(comp.size1$emmeans$Yeast, 
                               comp.size1$emmeans$DAI,
                               str_trim(comp.size1$emmeans$.group))
colnames(sig.diff.letters) <- c("Yeast", 
                                "DAI",
                                "Letters")

# for plotting with letters from significance test
size1plot <- filterdata1 %>%
  group_by(Yeast, DAI) %>%
  dplyr::summarize(
    colonydiameter = mean(increase, na.rm=TRUE),
    se = sd(increase)/sqrt(3)) %>%
  left_join(sig.diff.letters) 


colonysize1 <- filterdata1 %>%
  ggplot(aes(x = Yeast, y = increase, group = Yeast, fill = Yeast)) +
  geom_boxplot()+
  geom_point(shape = 21, color = "black", position = position_jitterdodge(dodge.width = 0.9))+
   geom_text(data = size1plot, aes(label = Letters, y = colonydiameter+(3*se)), vjust = -0.75, hjust = 2) +
  ylab("Increase in Colony Diameter (mm)") +   
  xlab("Yeast Isolates") +  
    scale_fill_manual(values = cbbPalette) +
  theme(strip.text.x = element_text(size = 18, face= "italic"))+
  theme(axis.title.x = element_text(size =25, face = "bold"), axis.title.y = element_text(size = 25, face = "bold"), axis.text.x=element_text(size=20, angle= 0), axis.text.y=element_text(size=20, angle= 0))+
  theme_classic() +
  theme(legend.position = "none")+
  ggtitle(expression(paste("Impact of Yeast on Growth of ", italic("Methylobacterium platani"), " EMM_B52 (Biological Replicate 1)"))) +
   theme(plot.title = element_text(face = "bold", color = "Blue2", size = 12, hjust = 0.5))+
  facet_wrap(~DAI)
colonysize1
```

### Impact of Yeast on Colony Weight

```{r Colony Weight Plot for Biological Replicate 1}
#filter data to avoid other data except of last day since it was detructive sample taking procedure and colony was taken out from the plate to measure the weight on the last day of data collection.

filterdata1_2 <- filterdata1 %>% 
  filter(DAI == "8" )

#stats used for colony weight
cw1=lme(colony_weight~Yeast,data = filterdata1_2, random = ~1|Replication) ##replication random effect added by Dr. Steury
summary(cw1)
Anova(cw1)
#pairwise comparision
lsmeans.cw1 <- emmeans(cw1, ~ Yeast) 

# contrast with Tukey ajustment 
comp.cw1 <- multcomp::cld(object = lsmeans.cw1, alpha = 0.05, Letters = letters, reversed = TRUE, details = TRUE)
comp.cw1

# Extracting the letters for the bars
sig.diff.letters <- data.frame(comp.cw1$emmeans$Yeast, 
                               str_trim(comp.cw1$emmeans$.group))

colnames(sig.diff.letters) <- c("Yeast", "Letters")

# for plotting with letters from significance test
weightplot <- filterdata1_2 %>%
  group_by(Yeast) %>%
  dplyr::summarize(
    colonyweight = mean(colony_weight, na.rm=TRUE),
    se = sd(colony_weight)/sqrt(3)) %>%
  left_join(sig.diff.letters) 

 colonyweightplot <- filterdata1_2 %>%
  ggplot(aes(x = Yeast, y = colony_weight, group = Yeast, fill = Yeast)) +
  geom_boxplot()+
  geom_point(shape = 21, color = "black", position = position_jitterdodge(dodge.width = 0.9))+
  geom_text(data = weightplot, aes(label = Letters, y = colonyweight+(2*se)), vjust = -0.5, hjust = 2) +
  ylab("Colony Weight (gram)") +   
  xlab("Yeast Isolates") +  
  scale_fill_manual(values = cbbPalette) +
  theme(strip.text.x = element_text(size = 18, face= "italic"))+
  theme(axis.title.x = element_text(size =25, face = "bold"), axis.title.y = element_text(size = 25, face = "bold"), axis.text.x=element_text(size=20, angle= 0), axis.text.y=element_text(size=20, angle= 0))+
  theme_classic() +
  theme(legend.position = "none")
  #ggtitle(expression(paste("Impact of Yeast on Growth of ", italic("Methylobacterium platani"), " EMM_B52"))) +
   #theme(plot.title = element_text(face = "bold", color = "Blue2", size = 12, hjust = 0.5))
colonyweightplot
```


### Impact of Yeast on Colony Forming Unit (CFU) 

```{r Colony Forming Unit(CFU) Plot for Biological Replicate 1}
#stats used for colony forming unit
cfu1=lme(log10(CFU)~Yeast,data = filterdata1_2, random = ~1|Replication, na.action = na.omit) 
summary(cfu1)
Anova(cfu1)
#pairwise comparision
lsmeans.cfu1 <- emmeans(cfu1, ~ Yeast) ###interaction added by Dr. Steury
#emmeans(cfu,pairwise~Yeast) ##added by Dr. steury
# estimate lsmeans 
comp.cfu1 <- multcomp::cld(object = lsmeans.cfu1, alpha = 0.05, Letters = letters, reversed = TRUE, details = TRUE)
# contrast with Tukey ajustment 
comp.cfu1


# Extracting the letters for the bars
sig.diff.letters <- data.frame(comp.cfu1$emmeans$Yeast, 
                               str_trim(comp.cfu1$emmeans$.group))

colnames(sig.diff.letters) <- c("Yeast", "Letters")

# for plotting with letters from significance test
cfu1plot <- filterdata1_2 %>%
  filter(Replication %in% c(1:3)) %>% 
  group_by(Yeast) %>%
  dplyr::summarize(
    cfu1 = mean(log10(CFU), na.rm=TRUE),
    se = sd(log10(CFU))/sqrt(3)) %>%
  left_join(sig.diff.letters) 

 
CFU1PLOT <- filterdata1 %>%
  subset(Media=="YePDA") %>%
  #subset(Yeast %in% c("Control", "SP_F14","EMM_F3", "EMM_F34")) %>%
  ggplot(aes(x = Yeast, y = log10(CFU), group = Yeast, fill = Yeast)) +
  geom_boxplot()+
  geom_point(shape = 21, color = "black", position = position_jitterdodge(dodge.width = 0.9))+
  geom_text(data = cfu1plot, aes(label = Letters, y = cfu1+(4*se)), vjust = -0.1, hjust = 2) +
  ylab("log of Colony Forming Unit per ml") +   
  xlab("Yeast Isolates") +  
  scale_fill_manual(values = cbbPalette) +
  theme(strip.text.x = element_text(size = 18, face= "italic"))+
  theme(axis.title.x = element_text(size =25, face = "bold"), axis.title.y = element_text(size = 25, face = "bold"), axis.text.x=element_text(size=10, angle= 0), axis.text.y=element_text(size=20, angle= 0))+
  theme_classic() +
  theme(legend.position = "none")
  #ggtitle("Impact of Yeast on Growth of Methylobacterium platani EMM_B52") +
  #theme(plot.title = element_text(hjust = 0.5))
CFU1PLOT
```


```{r CFU and colony weight Combined Plot for Biological Replicate 1}
#combining similar dataset plots into one
cw1andcfu1 <- ggarrange(colonyweightplot, CFU1PLOT, nrow = 1, ncol = 2, common.legend = F)
  
cw1andcfu1 <- annotate_figure(cw1andcfu1, top = text_grob(expression(paste("Impact of Yeast on Growth of ", italic("Methylobacterium platani"), " EMM_B52 (Biological  Replicate 1)")), color = "Blue2", size = 12, face = "bold", hjust = 0.5))
cw1andcfu1
```

## Biological Replicate 2

```{r Load data for Biological Replicate 2}
#read data
BioRep2 <- read.csv("Splitplate/SplitPlateData/2024-11-11_SplitPlate_B52-1.csv", na.strings = "na")
#load the color palettes
cbbPalette <- c( "#E69F00", "#56B4E9", "#009E73", "#CC79A7")
#check the structure and see the data
str(BioRep2)
head(BioRep2)
#changing the variable into categorical data
BioRep2$Replication=as.factor(BioRep2$Replication)
BioRep2$Yeast=as.factor(BioRep2$Yeast)
BioRep2$DAI=as.factor(BioRep2$DAI)
```

### Impact of Yeast on Colony Size 

```{r Increase in colony size Plot for Biological Replicate 2}
#removing the 1st data since there was no growth associated with it
#the data is considered to be the baseline for growth so since it violates the assumption of normality, we remove it from the analysis. For this Day 10 data was also collected. Thus, for consistency, we remove the data for visualization and analysis.
filterdata2 <-BioRep2 %>% 
filter(DAI != "2", DAI != "10")

#load library
library(nlme)
#stats used for size
size2=lme(increase~Yeast*DAI,data = filterdata2,random=~1|Replication) ##replication random effect added by Dr. Steury
summary(size2)
anova(size2)
#pairwise comparision
lsmeans.size2 <- emmeans(size2, ~ Yeast|DAI) ###interaction added by Dr. Steury
#emmeans(size,pairwise~Yeast) ##added by Dr. steury
# estimate lsmeans 
comp.size2 <- multcomp::cld(object = lsmeans.size2, alpha = 0.05, Letters = letters, reversed = TRUE, details = TRUE)
# contrast with Tukey ajustment 
comp.size2


# Extracting the letters for the bars
sig.diff.letters <- data.frame(comp.size2$emmeans$Yeast, 
                               comp.size2$emmeans$DAI,
                               str_trim(comp.size2$emmeans$.group))
colnames(sig.diff.letters) <- c("Yeast", 
                                "DAI",
                                "Letters")

# for plotting with letters from significance test
size2plot <- filterdata2 %>%
  group_by(Yeast, DAI) %>%
  dplyr::summarize(
    colonydiameter = mean(increase, na.rm=TRUE),
    se = sd(increase)/sqrt(3)) %>%
  left_join(sig.diff.letters) 


colonysize2 <- filterdata2 %>%
  ggplot(aes(x = Yeast, y = increase, group = Yeast, fill = Yeast)) +
  geom_boxplot()+
  geom_point(shape = 21, color = "black", position = position_jitterdodge(dodge.width = 0.9))+
   geom_text(data = size2plot, aes(label = Letters, y = colonydiameter+(3*se)), vjust = -0.75, hjust = 2) +
  ylab("Increase in Colony Diameter (mm)") +   
  xlab("Yeast Isolates") +  
    scale_fill_manual(values = cbbPalette) +
  theme(strip.text.x = element_text(size = 18, face= "italic"))+
  theme(axis.title.x = element_text(size =25, face = "bold"), axis.title.y = element_text(size = 25, face = "bold"), axis.text.x=element_text(size=20, angle= 0), axis.text.y=element_text(size=20, angle= 0))+
  theme_classic() +
  theme(legend.position = "none")+
  ggtitle(expression(paste("Impact of Yeast on Growth of ", italic("Methylobacterium platani"), " EMM_B52 (Biological Replicate 2)"))) +
   theme(plot.title = element_text(face = "bold", color = "Blue2", size = 12, hjust = 0.5))+
  facet_wrap(~DAI)
colonysize2
```

### Impact of Yeast on Colony Weight

```{r Colony weight plot for Biological Replicate 2}
#filter data to avoid other data except 
filterdata2_2 <- BioRep2 %>% 
  filter(DAI == "10" )

#stats used for colony weight
cw2=lme(colony_weight~Yeast,data = filterdata2_2, random = ~1|Replication) 
summary(cw2)
Anova(cw2)
#pairwise comparision
lsmeans.cw2 <- emmeans(cw2, ~ Yeast) 

# contrast with Tukey ajustment 
comp.cw2 <- multcomp::cld(object = lsmeans.cw2, alpha = 0.05, Letters = letters, reversed = TRUE, details = TRUE)

comp.cw2

# Extracting the letters for the bars
sig.diff.letters <- data.frame(comp.cw2$emmeans$Yeast, 
                               str_trim(comp.cw2$emmeans$.group))

colnames(sig.diff.letters) <- c("Yeast", "Letters")

# for plotting with letters from significance test
weightplot <- filterdata2_2 %>%
  group_by(Yeast) %>%
  dplyr::summarize(
    colonyweight = mean(colony_weight, na.rm=TRUE),
    se = sd(colony_weight)/sqrt(3)) %>%
  left_join(sig.diff.letters) 

 colonyweightplot2 <- filterdata2_2 %>%
  ggplot(aes(x = Yeast, y = colony_weight, group = Yeast, fill = Yeast)) +
  geom_boxplot()+
  geom_point(shape = 21, color = "black", position = position_jitterdodge(dodge.width = 0.9))+
  geom_text(data = weightplot, aes(label = Letters, y = colonyweight+(2*se)), vjust = -0.5, hjust = 2) +
  ylab("Colony Weight (gram)") +   
  xlab("Yeast Isolates") +  
  scale_fill_manual(values = cbbPalette) +
  theme(strip.text.x = element_text(size = 18, face= "italic"))+
  theme(axis.title.x = element_text(size =25, face = "bold"), axis.title.y = element_text(size = 25, face = "bold"), axis.text.x=element_text(size=20, angle= 0), axis.text.y=element_text(size=20, angle= 0))+
  theme_classic() +
  theme(legend.position = "none")
colonyweightplot2
```

### Impact of Yeast on Colony Forming Unit (CFU)

```{r Colony Forming Unit(CFU) Plot for Biological Replicate 2}
#stats used for colony forming unit
cfu2=lme(log10(CFU)~Yeast,data = filterdata2_2, random = ~1|Replication, na.action = na.omit) 
summary(cfu2)
Anova(cfu2)
#pairwise comparision
lsmeans.cfu2 <- emmeans(cfu2, ~ Yeast) ###interaction added by Dr. Steury
#emmeans(cfu,pairwise~Yeast) ##added by Dr. steury
# estimate lsmeans 
comp.cfu2 <- multcomp::cld(object = lsmeans.cfu2, alpha = 0.05, Letters = letters, reversed = TRUE, details = TRUE)
# contrast with Tukey ajustment 
comp.cfu2


# Extracting the letters for the bars
sig.diff.letters <- data.frame(comp.cfu2$emmeans$Yeast, 
                               str_trim(comp.cfu2$emmeans$.group))

colnames(sig.diff.letters) <- c("Yeast", "Letters")

# for plotting with letters from significance test
cfu2plot <- filterdata2_2 %>%
  na.omit() %>% 
  group_by(Yeast) %>%
  dplyr::summarize(
    cfu2 = mean(log10(CFU), na.rm=TRUE),
    se = sd(log10(CFU))/sqrt(3)) %>%
  left_join(sig.diff.letters) 

 
CFU2PLOT <- filterdata2_2 %>%
  ggplot(aes(x = Yeast, y = log10(CFU), group = Yeast, fill = Yeast)) +
  geom_boxplot()+
  geom_point(shape = 21, color = "black", position = position_jitterdodge(dodge.width = 0.9))+
  geom_text(data = cfu2plot, aes(label = Letters, y = cfu2+(4*se)), vjust = -0.1, hjust = 2) +
  ylab("log of Colony Forming Unit per ml") +   
  xlab("Yeast Isolates") +  
  scale_fill_manual(values = cbbPalette) +
  theme(strip.text.x = element_text(size = 18, face= "italic"))+
  theme(axis.title.x = element_text(size =25, face = "bold"), axis.title.y = element_text(size = 25, face = "bold"), axis.text.x=element_text(size=10, angle= 0), axis.text.y=element_text(size=20, angle= 0))+
  theme_classic() +
  theme(legend.position = "none")
CFU2PLOT
```

```{r CFU and Colony weight Combined Plot for Biological Replicate 2}
#combining similar dataset plots into one
cw2andcfu2 <- ggarrange(colonyweightplot2, CFU2PLOT, nrow = 1, ncol = 2, common.legend = F)
  
cw2andcfu2 <- annotate_figure(cw2andcfu2, top = text_grob(expression(paste("Impact of Yeast on Growth of ", italic("Methylobacterium platani"), " EMM_B52 (Biological Replicate 2)")), color = "Blue2", size = 12, face = "bold", hjust = 0.5))
cw2andcfu2
```

